---
title: "RegexExercises"
output: html_document
---

##Exercises


1. Find a pattern that matches the following rows:

abcdefg

abcde

abc

```{r, echo=FALSE}
a <- c("abcdefg","abcde","abc")
```



2. Write a pattern that matches all the digits in the strings. Anywhere within the string:

abc123xyz	

define "123"	To be completed

var g = 123;

```{r, echo=FALSE}
b <- c("abc123xyz","define 123","abcvar g = 123;")
```



3. Write a pattern that matches the first three strings, but not the last. 

cat.

896.	

?=+.	

abc1
```{r, echo=FALSE}
c <- c("cat.","896.","?=+.","abc1")
```



4. Match the first three strings, but not the last three strings

can

man

fan

dan

ran

pan

```{r, echo=FALSE}
d <- c("can","man","fan","dan", "ran", "pan")
```



5.Use the bracket notation to match the first three lines (all them have a pattern)

Ana

Bob

Cpc

aax

bby

ccz

```{r, echo=FALSE}
e <- c("Ana","Bob","Cpc","aax", "bby", "ccz")
```



6. Try writing a pattern that matches only the first two spellings

wazzzzzup

wazzzup

wazup

```{r, echo=FALSE}
f <- c("wazzzzzup","wazzzup","wazup")
```


7. Skip the last string

aaaabcc

aabbbbc

aacc

a

```{r, echo=FALSE}
g <- c("aaaabcc","aabbbbc","aacc","a")
```


8. Notice how the plurality of the word "file" depends on the number of files found. Try writing a pattern to match only the lines where one or more files were found

1 file found

2 files found

24 files found

No files found

```{r, echo=FALSE}
h <- c("1 file found","2 files found","24 files found","No files found")
```


9. Try writing a pattern that can match each line regardless of how much whitespace is between the number and the content

 abc

    abc

       abc

abc

```{r, echo=FALSE}
i <- c("1. abc","2.  abc","3.    abc","4.abc")
```



10. Match the line where only the word successful appears

Mission: successful

Last Mission: unsuccessful

Next Mission: successful upon capture of targe

```{r, echo=FALSE}
j <- c("Mission: successful","Last Mission: unsuccessful","Next Mission: successful upon capture of target")
```


11. Write a regular expression that matches only the filenames (not including extension) of the PDF's below

file_record_transcript.pdf

file_0724249.pdf

testfile_fake.pdf.tmp

```{r, echo=FALSE}
k <- c("file_record_transcript.pdf","file_0724249.pdf","testfile_fake.pdf.tmp")
```


12.Write an expression that matches and captures both the full date, as well as the year of the date

Jan 1987

May 1969

Aug 2011

```{r, echo=FALSE}
l <- c("Jan 1987","May 1969","Aug 2011")
```


13. Try to capture the width and height of each display

1289x720

1920x1600

1024x768

```{r, echo=FALSE}
m <- c("1289x720","1920x1600","1024x768")
```


14. Write a conditional pattern that matches only the lines with small fuzzy createrues below

I love cats

I love dogs

I love logs

I love cogs

```{r, echo=FALSE}
n <- c("I love cats","I love dogs","I love logs","I love cogs")
```


##Answers

 Question 1
```{r}
grep("[a-z]",a, ignore.case=TRUE)
```

Question 2
```{r}
grep("[0-9]",b)
#or
grep("\\d",b)
```

Question 3
```{r}
grep("^.*\\.$",c, ignore.case=TRUE)
```

Question 4 
```{r}
grep("[^drp]an",d, ignore.case=TRUE)
```

Question 5
```{r}
grep("[A-Z][a-z][a-z]",e)
```

Question 6
```{r}
grep("z{2,}",f)
#or
grep("waz{3,5}up",f)
```

Question 7 
```{r}
grep("aa+b*c+",g)
#or
grep("a{2,4}b{0,4}c{1,2}",g)
```

Question 8
```{r}
grep("[0-9]+\\sfiles?\\sfound?",h)
#or
grep("\\d+ files? found?",h)
```

Question 9 
```{r}
grep("\\s+",i)
```

Question 10
```{r}
grep("[^un]successful$",j)
```

Question 11
```{r}
grep("^(file.+)(.pdf)$",k)
```

Question 12 
```{r}
grep("(\\w+\\s+(\\d+))",l)
```

Question 13 
```{r}
grep("(\\d+)x(\\d+)",m)
```

Question 14
```{r}
grep("I love (cats|dogs)",n)
```