---
title: "Regex Introduction"
output: html_document
---

#Regex in R

##Introduction

Whenever you want to extract information from text, regular expressions are extremely useful. 
R has some functions to write patterns to match specific sequence of characters. 


Using Regex functions, we can:

1. Search matches

2. Extract parts of a string

3. Find and replace characters


##1. Search matches 


In R, exist some functions to search patters in text strings

- grep: returns a vector of the indices of the elements of x that yielded a match

- grepl: returns a logical vector of the indices of the elements of x that yielded a match

- regexpr: returns an integer vector of the same length as text giving the starting position of the first match  or -1 if there is none, with attribute "match.length", an integer vector giving the length of the matched text (or -1 for no match)


We are going to use the following vector containing text in order to show some examples of Regex usage:

```{r}
#put words in a vector to practice some Regex
x <- c("16_24cat", "25_34cat", "35_44catch", "45_54Cat", "55_104fat","cat_98","78_cat","hello.txt", "The", "licenses", "for", "most", "software", "are",   "designed", "to", "take", "away", "your", "freedom",   "to", "share", "and", "change", "it.",    "", "By", "contrast,", "the", "GNU", "General", "Public", "License", "is", "intended", "to", "guarantee", "your", "freedom", "to",
   "share", "and", "change", "free", "software", "--", "to", "make", "sure", "the", "software", "is", "free", "for", "all", "its", "users", "english muffin", "Sara.txt", "IMG23.png", "panchita45@hotmail.com")
```



#### Select all the items in strings containing a word (case sensitive)

We want to find match all the words containing the word "cat"
```{r}
grep(pattern = "cat", x = x)
```

#### Select all the items containing a word at the end of a string (not case sensitive)(greedy)

Match all the words that contain the word "cat" at the end of the string
```{r}
grep("cat$", x, ignore.case = T)
```

### Select all the items containing a word at the beginning of a string (case sensitive)

Match all the words that contain the word "cat" at the beginning of the string
```{r}
grep("^[cat].*", x)
```

### Select specific file extension

Sometimes we just need to extract the documents with certain file extensions. E.g. .txt, .pdf or .ppt files. In this case, we will match all the .txt files
```{r}
grep("^.*\\.txt$", x)    
```

We can select the full name of a file
```{r}
grep("^(IMG\\d+\\.png)$", x)
```

Or just capture and extract the file name without the extension (Capture the part before the period)
```{r}
grep("^(IMG\\d+)\\.png$", x)
```




##2. Extract parts of a string


The function *strstring* helps to split the elements of a character vector into substrings according the matches we want to find. 

####Select all the characters eliminating certain symbol 

In this example, we elmininate the symbol "_" and extract the words left

```{r}
strsplit(x, split = "_")
```



##3. Find and replace

The most used expresions to find a pattern and replace it with something are:

- sub
- gsub

We want all instances of two numbers followed by an alphabet letter to have an additional character
inserted, but only if the first of those numbers is 3 or less (case sensitive)
```{r}
gsub("([1-3][1-9][a-z])", "m\\1", x, perl = T)
#[1-3] numbers of 3 or less
#[1-9] instances of numbers
#[a-z] followed by an alphabet letter
#\\1 replace with the value captured in group 1
#m add "m" in the pattern
```


